# Software

## Installation

R is open source statistical software that can be downloaded from [CRAN](https://cran.r-project.org/).

Once R has been installed the two R packages associated with this course can be installed by running the following code in the R console.
```{r, eval=FALSE}
install.packages("pak")
pak::pak("poissonconsulting/DFOsalmonids")
pak::pak("poissonconsulting/DFOsamplestats")
```

The packages can then be loaded using

```{r}
library(DFOsalmonids)
library(DFOsamplestats)
```

## Data Package

The `DFOsalmonids` data package provides two data sets both provided by the Department of Fisheries and Oceans.

The first is catch and release data.

```{r}
library(tibble)

DFOsalmonids::salmonids
```

The second scale age data.

```{r}
DFOsalmonids::scale_age
```

## Analytic Package

[`DFOsamplestats`](https://poissonconsulting.github.io/DFOsamplestats/index.html) provides functions to calculate the power (at a significance level of 0.05) using [`rate2_power()`](https://poissonconsulting.github.io/DFOsamplestats/reference/rate2_power.html) and sample size to achieve a power of 0.8 for the difference in the rate between two groups using [`rate2_samplesize()`](https://poissonconsulting.github.io/DFOsamplestats/reference/rate2_samplesize.html).

```{r}
library(DFOsamplestats)
rate2_power(p1 = 0.005, p2 = 0.0075, n = 31197)
rate2_samplesize(p1 = 0.005, p2 = 0.0075)
```

`DFOsamplestats` also provides the [`rate_effect()`](https://poissonconsulting.github.io/DFOsamplestats/reference/rate_effect.html) to estimate the effect size and p-values  based on the number of individuals returning in each group given the number tagged.

```{r}
rate_effect(r = c(156,  234), n = 31197)
```

The p-value for the first group is for $p_1 = 0.5$ while the p-value for subsequent groups is for $p_n = p_1$.

And the [`rate_sim()`](https://poissonconsulting.github.io/DFOsamplestats/reference/rate_sim.html) simulate data.

```{r}
set.seed(99)
rate_sim(p = c(0.005, 0.0075), n = 31197)
rate_sim(p = c(0.005, 0.0075), n = 31197)
```

`DFOsamplestats` also provides the [`rate2_power_analysis()`](https://poissonconsulting.github.io/DFOsamplestats/reference/rate2_power_analysis.html) function to perform a full power analysis.

```{r}
set.seed(100)
rate2_power_analysis(p1 = 0.005, p2 = 0.0075, n = 31197)
```

### Bayesian

Finally `DFOsamplestats` also provide [`rate_effect_bayesian()`](https://poissonconsulting.github.io/DFOsamplestats/reference/rate_effect_bayesian.html) and [`rate2_power_analysis_bayesian()`](https://poissonconsulting.github.io/DFOsamplestats/reference/rate2_power_analysis_bayesian.html) to estimate effect sizes and power incorporating prior information.

```{r}
set.seed(100)
rate_effect_bayesian(r = 1, n = 10, alpha = 5, beta = 5)
rate2_power_analysis(p1 = 0.005, p2 = 0.0075, n = 31197, nsims = 100)
```

### Heuristics

It is worth noting that `rate2_power()`, which is a wrapper on `stats::power.prop.test()`,
uses a heuristic to calculate the power.
As a result it is quicker but less reliable at small sample sizes than `rate2_power_analysis()` which performs a full power analysis.
However, `rate2_power_analysis()` still assumes that the likelihood profile is normally distributed which is typically not the case with little data at very low or high rates.
For the most reliable estimate use `rate2_power_analysis_bayesian()` with `nsims = 1000`.

```{r}
set.seed(100)
rate_effect(0, 5)
rate_effect_bayesian(0, 5)

rate2_power(0.1, 5)
rate2_power_analysis(0.1, 5)
rate2_power_analysis_bayesian(0.1, 5, nsims = 100)
```
