# Scenarios

Before considering the scenarios it is important to load the `DFOsamplestats` package.

```{r}
library(DFOsamplestats)
```

## Scenario 1 (Fish Tagging)

Consider two groups of fish. Group 1 is standard release timing and Group 2 is late release timing.
Historic data suggests that fish in Group 1 have return rate of 0.5% while fish in Group 2 have a return rate of 0.75%.

> Question: How many samples do we need to release in total to be able to detect this expected difference in survival?

We can answer this question using `DFOsamplestats::rate2_samplesize()`.

```{r}
rate2_samplesize(p1 = 0.005, p2 = 0.0075, alternative = "greater")
```
This is the total number of fish that should be tagged (split equally between the two groups).

We can confirm that we have a power of 0.8 using `rate2_power()`.

```{r}
rate2_power(p1 = 0.005, n = 24574, p2 = 0.0075, alternative = "greater")
```

It is worth noting that the slower but more reliable `rate2_power_analysis()` gives a very similar answer.

```{r}
set.seed(99)
rate2_power_analysis(p1 = 0.005, n = 24574, p2 = 0.0075, alternative = "greater")
```

The `rate2_power_analysis()` which performs a full power analysis by simulating data can also be used to estimate the power to detect a difference for one group relative to 0.5

## Scenario 2 (Scale Aging)

Fish return at age 2, 3, 4, and 5. Age 5 fish comprise about 10% of the returns based on historic scale data we have. 

> Question: How many scales samples do we need to collect in total to be able to detect a 2% difference the age 5 component?

```{r}
rate2_samplesize(p1 = 0.1, p2 = 0.12)
```

This is the total number of scale samples that should be taken (split evenly between the two years) if we are unwilling to assume the direction of change.

```{r}
rate2_power(p1 = 0.1, p2 = 0.12, n = 7682)
```

If we were willing to assume that any change was positive then we would have greater power.

```{r}
rate2_power(p1 = 0.1, p2 = 0.12, n = 7682, alternative = "greater")
```
  
### Scenario 3 (Age Proportions)

Fish return at age 2 (5%), 3 (30%), 4 (50%), and 5 (15%).

A useful way to understand the uncertainty at different sample sizes is to simulate a data set and estimate the values.

```{r}
set.seed(42)
sim <- rate_sim(p = c(0.05, 0.3, 0.5, 0.15), n = 1000)
effect <- rate_effect(r = sim$r, n = sim$n)
```
It is worth noting that the first p-value is the null hypothesis that p1 = 0.5 while the subsequent p-values are for the null hypothesis that pn = p1.

The estimated rates and associated 95% CIs can be plotted using the `ggplot2` package.

```{r}
library(ggplot2)
library(scales)

ggplot(data = effect) +
  aes(x = group, y = estimate) +
  geom_pointrange(aes(ymin = lower, ymax = upper)) +
  expand_limits(y = 0) +
  scale_x_discrete("Group", labels = paste0("Age-", 2:5)) +
  scale_y_continuous("Percent (%)", labels = percent)
```



<!-- > Questions: How many scale (age) samples do we need to collect to confidently assert that our sample is representative of the true population age proportions? -->

<!-- Does the size of the population affect this (i.e. does the sample size different when one population sees an average return of 500 vs 10,000? -->

<!-- Can we pool age results across multiple years? Can we simply take an average? What are the considerations/caveats to this approach -->

<!-- Is the sample size smaller if there are less age classes? How do we factor this into sample size calculation? -->

