# Scenarios

Before considering the scenarios it is important to load the `DFOsamplestats` package.

```{r}
library(DFOsamplestats)
```

## Fish Tagging Two Groups

Consider two groups of fish. Group 1 is standard release timing and Group 2 is late release timing.
Historic data suggests that fish in Group 1 have return rate of 0.5% while fish in Group 2 have a return rate of 0.75%.

> Question: How many fish do we need to tag in total to be able to detect this expected difference in survival?

We can answer this question using `rate2_samplesize()`.

```{r}
rate2_samplesize(p1 = 0.005, p2 = 0.0075)
```
This is the total number of fish that should be tagged (split equally between the two groups).

If we are willing to assume that any difference is positive then the sample size would be smaller.

```{r}
rate2_samplesize(p1 = 0.005, p2 = 0.0075, alternative = "greater")
```

We can confirm that we have a power of 0.8 using `rate2_power()`.

```{r}
rate2_power(p1 = 0.005, n = 24574, p2 = 0.0075, alternative = "greater")
```
As the data set is relatively big the slower but more reliable `rate2_power_analysis()` gives a very similar answer.

```{r}
set.seed(99)
rate2_power_analysis(p1 = 0.005, n = 24574, p2 = 0.0075, alternative = "greater")
```

## Fish Tagging Multiple Groups

Consider the following data set

```{r, echo=FALSE}
data <- tibble::tribble(
  ~year, ~tagged, ~catch, ~escapement,
2017, 10000, 1291,	152,
2018, 10000,	1465,	210,
2019, 10000, 898,	165,
2020, 10000,	1876,	256)
print(data)

```

> Question: Are the rates significantly different across years / which years? How do I create confidence intervals?

This can be done using `rate_effect()`. 
It is worth noting that the first p-value is the null hypothesis that $p_1 = 0.5$ while the subsequent p-values are for the null hypothesis that $p_n = p_1$.

```{r}
effect <- rate_effect(r = data$escapement, n = data$tagged)
effect$year <- data$year
print(effect)
```

And the results plotted using `ggplot2`.

```{r}
library(ggplot2)
library(scales)

ggplot(data = effect) +
  aes(x = year, y = estimate) +
  geom_pointrange(aes(ymin = lower, ymax = upper)) +
  expand_limits(y = 0) +
  scale_x_continuous("Year") +
  scale_y_continuous("Percent (%)", labels = percent)
```

> Question: If not significant, how do I figure out how many more tags I should have applied / fisheries recoveries should I have made?

```{r}
p1 <- effect$estimate[1]
p2 <- effect$estimate[3]
rate2_samplesize(p1, p2)
```

## Age Proportions Two Groups

Fish return at age 2, 3, 4, and 5. 
Age 5 fish comprise about 10% of the returns based on historic scale data we have. 

> Question: How many scales samples do we need to collect in total to be able to detect a 2% difference in the age 5 component?

```{r}
rate2_samplesize(p1 = 0.1, p2 = 0.12)
```

This is the total number of scale samples that should be taken (split evenly between the two years) if we are unwilling to assume the direction of change.

```{r}
rate2_power(p1 = 0.1, p2 = 0.12, n = 7682)
```

If we were willing to assume that any change was positive then we would have greater power.

```{r}
rate2_power(p1 = 0.1, p2 = 0.12, n = 7682, alternative = "greater")
```
  
## Age Proportions Multiple Groups

Fish return at age 2 (5%), 3 (30%), 4 (50%), and 5 (15%).

A useful way to understand the uncertainty at different sample sizes is to simulate a data set and estimate the values.

```{r}
set.seed(42)
sim <- rate_sim(p = c(0.05, 0.3, 0.5, 0.15), n = 1000)
effect <- rate_effect(r = sim$r, n = sim$n)
```

The estimated rates and associated 95% CIs can be plotted using the `ggplot2` package.

```{r}
ggplot(data = effect) +
  aes(x = group, y = estimate) +
  geom_pointrange(aes(ymin = lower, ymax = upper)) +
  expand_limits(y = 0) +
  scale_x_discrete("Group", labels = paste0("Age-", 2:5)) +
  scale_y_continuous("Percent (%)", labels = percent)
```



<!-- > Questions: How many scale (age) samples do we need to collect to confidently assert that our sample is representative of the true population age proportions? -->

<!-- Does the size of the population affect this (i.e. does the sample size different when one population sees an average return of 500 vs 10,000? -->

<!-- Can we pool age results across multiple years? Can we simply take an average? What are the considerations/caveats to this approach -->

<!-- Is the sample size smaller if there are less age classes? How do we factor this into sample size calculation? -->
