# Prior Information

Bayesian analysis allows the incorporation of prior information.
Incorporating existing knowledge represents a form of regularization in the sense that biologically implausible results are not considered reasonable without adequate support in the data.

## Five Males

Consider a case where a biologist captures five returning salmon at random from run of 10,000 individuals and discovers that they are all males.

### Frequentist with Exact Estimate

Under a strict interpretation of the frequentist framework they should tell their manager that the sex ratio is skewed towards males (the p-value for the null hypothesis of an equal sex ratio is `dbinom(0, 5, 0.5)` = 0.031 and that their best estimate is that all 10,000 individuals are males!

They could also publish a paper in a high impact journal that only publishes surprising results discussing their statistically significant finding.

### Frequentist with Approximate Estimate

If they were use a heuristic to calculate an approximate estimate they would still report that the best estimate is that all the fish are males although they wouldn't be able to rule out the possibility that all the fish are females!

```{r}
library(DFOsamplestats)
rate_effect(0, 5)
```

### Bayesian with Uninformative Prior

If they were fit the equivalent Bayesian model with an uninformative prior they would still conclude that the sex ratio was significantly different from 50:50 but their estimate (if using the median value of the posterior probability) would be that approximately 11% of the run was female.

```{r}
set.seed(42)
rate_effect_bayesian(0, 5)
```

### Bayesian with Weakly Informative Prior

If they were to fit the Bayesian model with a weakly informative prior say a beta distribution with alpha = 11 and beta = 11

```{r}
plot_dbeta(11, 11)
```

then they would estimate that about 40% of the run was female although they would consider it reasonably likely that females make up 60% of the run.

```{r}
set.seed(42)
rate_effect_bayesian(0, 5, alpha = 11, beta = 11)
```

### Bayesian with Informative Prior

With an informative prior

```{r}
plot_dbeta(101, 101)
```

then they would conclude that knowing the sex of five fish barely changes the the likely sex ratio.

```{r}
set.seed(42)
rate_effect_bayesian(0, 5, alpha = 101, beta = 101)
```

Indeed given this much prior information then they would require 68 or more males from a sample size of 100 fish to obtain a statistically significant result.

```{r}
set.seed(42)
rate_effect_bayesian(32, 100, alpha = 101, beta = 101)
```

## Bayesian Updating

It is worth noting that this is exactly the same result as if we had had no prior knowledge about the likely value of the sex ratio (i.e. population biology did not exist as a discipline) and sampled 132 males out of 300 fish.
```{r}
set.seed(42)
rate_effect_bayesian(132, 300, alpha = 1, beta = 1)
```

It is also the same result as if our prior knowledge about the sex ratio had the same information content as sampling 132 males out of 300 fish and we had yet to collect any data.
```{r}
set.seed(42)
rate_effect_bayesian(NA_real_, 0, alpha = 133, beta = 169)
```

## Suprisal

Surprisal (in the Bayesian framework) is how surprising (in terms of tossing x heads in a row on a fair coin) it would be to discover the that true effect is in the opposite direction to the estimated effect [@rafi_semantic_2020].

```{r}
p2svalue(c(1, 0.5, 0.25, 0.05, 0.01, 0.001, 0.0001))
dbinom(0, 5, 0.5)
p2svalue(dbinom(0, 5, 0.5))
```
