# Distributions

A handful of probability distributions can be used to implement a surprising number of different models.

## Continuous Distributions

Continuous distributions describe the probability density for any given value within a particular range.

### Uniform Distribution

As its name suggests the [uniform distribution](https://en.wikipedia.org/wiki/Uniform_distribution_(continuous)) has a constant density between a lower and upper bound. The uniform distribution can be useful for defining an uninformative prior.

```{r, fig.cap = "A uniform distribution with a lower bound of 0 and an upper bound of 1."}
data <- data.frame(x = seq(-0.1, 1.1, length.out = 1000))
data$y <- dunif(data$x)
gp <- ggplot(data, aes(x = x, y = y)) +
  geom_line() +
  expand_limits(y = 0) +
  scale_y_continuous("Density")
gp + scale_x_continuous(breaks = c(0, 0.5, 1))
```

### Normal Distributions

Normal distributions are the most common type of distribution.

#### Normal Distribution

A [normal distribution](https://en.wikipedia.org/wiki/Normal_distribution) is a bell-shaped curve.

```{r, fig.cap = "A (standard) normal distribution with a mean of 0 and a standard deviation of 1."}
data <- data.frame(x = seq(-3, 3, length.out = 50))
data$y <- dnorm(data$x)

gp %+% data
```

#### Log-Normal Distribution

The [log-normal distribution](https://en.wikipedia.org/wiki/Log-normal_distribution) is a normal distribution on a logarithmic scale which means that all the possible values are positive. It is useful for modeling sizes or densities of organisms.

```{r, fig.cap = "A log-normal distribution with a mean on the log scale of 0 and a standard deviation on the log scale of 1"}
data <- data.frame(x = seq(0, 5, length.out = 100))
data$y <- dlnorm(data$x)

gp %+% data
```

### Beta Distribution

The [beta distribution](https://en.wikipedia.org/wiki/Beta_distribution) is a continous distribution on the probability scale which means that all the possible values are between 0 and 1.

```{r, fig.cap = "A beta distribution with an alpha (shape1) parameter of 5 and a beta (shape2) parameter of 2."}
data <- data.frame(x = seq(0, 1, length.out = 100))
data$y <- dbeta(data$x, 5, 2)

gp %+% data
```

## Discrete Distributions

Discrete distributions describe the probability for each of the possible outcomes.

#### Bernoulli Distribution

The [Bernoulli distribution](https://en.wikipedia.org/wiki/Bernoulli_distribution) gives the probabilities for binary outcomes such tossing a coin or whether an individual survives.

```{r, fig.cap="A Bernoulli distribution where the probability of outcome 1 is 0.7"}
data <- data.frame(x = 0:1)
data$y <- dbinom(data$x, 1, 0.7)

gp <- ggplot(data, aes(x = x, y = y)) +
  geom_col() +
  expand_limits(y = 0) +
  scale_y_continuous("Probability")

gp + scale_x_continuous(breaks = 0:10)
```

### Binomial Distribution

The [binomial distribution](https://en.wikipedia.org/wiki/Binomial_distribution) is an generalization of the Bernoulli distribution to multiple trials such the number of heads when tossing five coins or the number of marked fish that are resighted.
It is useful for modeling survival or mark-recapture data.

```{r, fig.cap="A binomial distribution for five trials with a probability of success of 0.7."}
data <- data.frame(x = 0:5)
data$y <- dbinom(data$x, 5, 0.7)
gp %+% data + scale_x_continuous(breaks = 0:5)
```

### Poisson Distribution

The [Poisson distribution](https://en.wikipedia.org/wiki/Poisson_distribution) gives the probabilities for a number of independent events such as observing an individual during a survey.
It is useful for analysing count data.

```{r, fig.cap="A Poisson distribution where the mean number of events is 3."}
data <- data.frame(x = 0:10)
data$y <- dpois(data$x, 3)
gp %+% data + scale_x_continuous(breaks = 0:10)
```
